In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-29 paper/supplementary_main.tex
Updates from Overleaf

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$. \mk{We always keep $R^i_2$, $R^i_1$ fixed in time in A.1.1 and A.1.4, right?}

@@

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2024modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation*}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation*}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
such that
\begin{equation*}
  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation*}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$, see also~\cite{masri2024discontinuous}.
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

@@

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure mass conservation, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, \mk{0}, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  \mk{n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
  \label{eq:def:sign}
\end{equation}
\mk{Why do we need the normal on non-IO vertices?}

@@

\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).

>>>>

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$, and remark that we interpret $R^i_2$ and $R^i_1$ as temporal averages (fixed in time), as the pulsations considered in \Cref{sec:app:peristalsis} are beyond the temporal resolution of our model. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$.

@@

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2024modelling, laurino2019derivation} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation*}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation*}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
such that
\begin{equation*}
  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation*}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$.
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

@@

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and for $y\in\mathcal{B}$ let $\mathcal{E}(y)$ denote the set of
(two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure mass conservation, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
  = 0.
  \label{eq:flux_conservation}
\end{equation}
Here the (normal ``vector") function $n^i$ takes the values in $\left\{-1, 1\right\}$ and 
defines an orientation of vertices of edge $\Lambda^i$. Specifically, for an edge $\Lambda^i$ oriented from $y_{\rm in}$ to $y_{\rm out}$ we set 
\begin{equation}
  n^i(y_{\rm in}) = 1,\quad
  n^i(y_{\rm{out}}) = -1. %\quad
  %\mk{n^i(y) = 0 \quad \foralls y \in 
  %\mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
\label{eq:def:sign}
\end{equation}
%\mk{Why do we need the normal on non-IO vertices?}

@@

\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
( \mu \nabla \bm u \cdot \bm{n} - p \bm n) \cdot \bm n  &= -R_0 ( \bm u \cdot \bm n ), \quad \bm u \cdot \bm t = 0
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm t = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ and $\bm t$ denote the unit outward normal and tangent vectors to the boundary respectively,  $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time, 
a discontinuous Galerkin (DG) finite element method with upwinding in space for the 3D domain to accurately capture sharp boundary layers, and a continuous Galerkin method for the 1D networks. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right) 
   + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  .
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{2\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$. The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$: 
\begin{align}\label{eq:1d_dg_disc}
\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.  
\end{align}
 In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}. 
 \begin{align}
    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
 h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
 \end{align}
 where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.
Note that the condition \eqref{eq:flux_conservation} is enforced weakly in the above formulation.
